---
title: "R coursework with Sumeet"
subtitle: "6th - Interactive Websites using R Shiny"
author: "Yura Song"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output:
  html_notebook:
    toc: yes
---

In the seventh and last session, we will learn how to use [R Shiny](https://shiny.rstudio.com/) to make interactive websites.

```{r setup, include=FALSE, cache=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = ".")
```

# Preparation

```{r libraries, message=FALSE}
## From Session 3
library(dplyr)
## From Session4
library(readxl)
library(tibble)
library(tidyr)
## From Session 5
library(ggplot2)
library(ggpubr)
library(coronavirus)
## From Session 6
library(shiny)
library(shinydashboard)
```

- Download [R Shiny Cheat Sheet](https://github.com/sumeetpalsingh/R_course/blob/master/CheatSheet/shiny.pdf) 

# R shiny

**R shiny** is the tool for sharing data with interactive websites! It would be helpful to present and share your data with your collaborators.

Actually you could work with the basic and default shiny app. You could run it by using **Run app** button. Also, before closing, you should stop the app with **stop** button.

## Component of R Shiny

R Shiny is composed of title, sidebar and main panel. Here, it does not have that much of flexibility so you should encode as a shiny feature. This is the minimum component of user interface. For instance,
```{r}
ui <- fluidPage(
  titlePanel("title"),
  
  sidebarLayout(
    sidebarPanel("sidebar"), ## User input
    mainPanel("main")
  )
)
```

## Input ID

We need to specify the input ID for specifying which one is input. This would be a link between input and output. It is called **inputId** and **outputId**.
```{r}
    # Sidebar with a slider input for number of bins 
    sidebarLayout(
        sidebarPanel(
            sliderInput(inputId = "bins",
                        "Number of bins:",
                        min = 1,
                        max = 50,
                        value = 30)
        ),

        # Show a plot of the generated distribution
        mainPanel(
           plotOutput(outputId = "distPlot")
        )
    )
```

## Server

It is always using pre-defined UI. Here, it will assign the output with the outputID on main panel. The ID is character vector and we need quotation mark for specification.
```{r}
# Define server logic required to draw a histogram
server <- function(input, output) {

    output$distPlot <- renderPlot({
        # generate bins based on input$bins from ui.R
        x    <- faithful[, 2] ## Taking second column : waiting time
        bins <- seq(min(x), max(x), length.out = input$bins + 1)

        # draw the histogram with the specified number of bins : Always last line is the output!!!
        hist(x, breaks = bins, col = 'darkgray', border = 'white')
    })
}
```

# Input and output

## Basic types of input

There are several different type of input and you could clarify the format of input with specific function.

- Checkbox, radio buttons
- Text, date input
- Numeric input, password input

## Output

### UI

We should clarify the format of output, such as plot, raw HTML, text or table.

### Server

Within the server, we should connect between input and output. *render()* creates the type of output we wish to make.