---
title: "Session6"
author: "Sumeet Pal Singh"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  html_document:
    toc: true
    toc_depth: 2
    toc_float:
      toc_collapse: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

***
***
Introduction to R Markdown
==========================

My first Markdown
-------------

### In R Language

#### Using RStudio

Today we will generate R Markdown text.  

This is useful for sharing R code.  

Plain-text can be *formatted* and remains **human readable**.

[Link to R Markdown Website](https://rmarkdown.rstudio.com/)

Tables can be made

Col1 | Col2 | Col3
---|---|---
1|2|3
54|34|5454

And lists

- Item 
- Another Item
  + Sub-item
  + Another Sub-item
    + Third layer

Or ordered

1. Item 1
2. Item 2

Let us start with R!

Include libraries for additional features

```{r libraries, message = FALSE, warning = FALSE}
## From Session 3
library(dplyr)
## From Session4
library(readxl)
library(tibble)
library(tidyr)
## From Session 5
library(ggplot2)
library(ggpubr)
library(coronavirus)
```

Let us look at the coronavirus dataset

```{r glimpse}
glimpse(coronavirus)

head(coronavirus)
```

Countries present in the dataset
Insert R code (Cmd + Option + I)
```{r countries}
table(coronavirus$Country.Region)
```



## Get Belgium Data

We will filter the data using dplyr for which Country.Region is Belgium
```{r}
coronavirus %>% 
  filter(Country.Region == "Belgium") -> 
  Belgium.covid
tail(Belgium.covid)
```


## Including Plots

You can also embed plots, for example:

```{r plot}
p.Belgium <- ggplot(data = Belgium.covid, 
                    mapping = aes(x = date, y = cases))
p.Belgium + geom_line(mapping = aes(color = type))

```

### Make smooth line graph

```{r}
## You don't need to initiate ggplot object again
p.Belgium 

p.Belgium + geom_smooth(mapping = aes(color = type, fill = type))

## Filling color manually

p.Belgium + 
  geom_smooth(mapping = aes(color = type, fill = type)) +
  scale_color_manual(values = c("blue", "cyan", "orange1")) +
  scale_fill_manual(values = c("blue", "cyan", "orange1"))
```

## Plot the ratio of dead to recovered

We will try to see the trend of death to recovery
```{r}
## Add a new column
Belgium.covid %>% 
  spread(key = type, value = cases) %>%
  group_by(date) %>%
  mutate(Dead_Recover = (death + 0.1) / (recovered + 0.1)) ->
  Belgium.covid
  
p.Belgium <- ggplot(data = Belgium.covid, 
                    mapping = aes(x = date, y = Dead_Recover))

p.Belgium + geom_smooth()
  
```


## Plot the same for UK and France



```{r}
sessionInfo()
```



