---
title: "R coursework with Sumeet"
subtitle: "3rd : Introduction to dplyr"
author: "Yura Song"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output:
  html_notebook:
    toc: yes
---

```{r setup, include=FALSE, cache=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = ".")
```

# Preparation

```{r libraries}
library(tidyverse)
```

We will keep working on **Tidy datasets**, composed of variables, observations and values.

# Dataset - mtcars

The dataset we will use today is called **mtcars**. This data was extracted from US motors magazine, composed of 32 observations on 11 numeric variables.
```{r}
data(mtcars)
```

## Getting glimpse of the data

```{r}
summary(mtcars)
```

Or we could use *View()* to see the dataframe as a format of table. Compared to *summary()*, it is much easier to go through the dataset and sorting the values are available. Also, you could filter the datasets as a part of datasets.
```{r}
View(mtcars)
```

Also, we could see the structure of each column with *str()*. This also provides a glimpse for the dataset.
```{r}
str(mtcars)
```

If we would like to know the size of dataset, we could use *dim()* function.
```{r}
dim(mtcars)
```

For the number of rows or columns, the function *nrow()* or *ncol()* will work!
```{r}
nrow(mtcars)
```

For getting the names of column or row, we could take them by using *colnames()* or *rownames()*.
```{r}
colnames(mtcars)
```

```{r}
rownames(mtcars)
```

Subsetting based on the rows is possible by giving the name of rows.
```{r}
mtcars[c("Mazda RX4", "Mazda RX4 Wag"), ]
```

## Correlation

To see the correlation between each elements, we could use *plot()*.
```{r}
plot(mtcars)
```

Also, if we would like to get the **correlation coefficient** between two different element in the dataset, we could use *cor()* function. Here, we should declare which element we would like to see. The default method is Pearson.
```{r}
cor(x=mtcars$mpg, y=mtcars$hp)
```

If we would like to use different method to calculate correlation coefficient, we could specify method.
```{r}
cor(x=mtcars$mpg, y=mtcars$hp, method="spearman")
```

If we give a subset of columns, it will return the correlation as a format of table.
```{r}
cor(mtcars[, c(1, 4)])
```

Also, correlation between all columns could be calculated too!
```{r}
cor(mtcars)
```

## Statistical test

### T-test

Does mpg differ based on Engine? We will see one by one. First, we will visualize data with **boxplot**.
```{r}
boxplot(formula = mpg ~ vs, data = mtcars)
```

Well, it looks quite different between each other. Here, to see their difference is significant or not, we will do *t.test()*.
```{r}
t.test(formula = mpg ~ vs, data = mtcars)
```

### ANOVA

Before starting analysis, we could make the element as a table. With function called *table()*, we could check the levels and number of observations.
```{r}
table(mtcars$gear)
```

Here, we would like to see whether the differences between gear is significant or not.

```{r}
boxplot(mpg~gear, data=mtcars)
```

Here, we will perform ANOVA. The function for ANOVA is *aov()*.
```{r}
mpg.model <- aov(formula = mpg ~ gear, data = mtcars)
summary(mpg.model)
```

It seems that we made mistake, the **Df** is a lot different.

```{r}
str(mtcars)
```

Well, gear is still as numeric value, not factor. Since the values on gear indicate the group, so we should give it as a **factor** in the model, not numeric values. If not, it will be considered as continuous values.
```{r}
mpg.model <- aov(formula = mpg ~ factor(gear), data = mtcars)
summary(mpg.model)
```

Here, also, we could see the pair-wise comparison. The model which we have used for ANOVA is given as input in *TukeyHSD()*.
```{r}
TukeyHSD(mpg.model)
```

## Exercise

Let's try to analyse **hp vs engine**.
```{r}
boxplot(hp ~ vs, data=mtcars)
```

```{r}
t.test(hp ~ vs, data=mtcars)
```

Then Let's try to analyse **hp vs carb**.
```{r}
table(mtcars$carb)
```

```{r}
hp.model <- aov(hp ~ factor(carb), data=mtcars)
summary(hp.model)
```

## Simple plotting

If we would like to see the relations between mpg and hp, we could visualise it by *plot()*. This function will decide which plot will be good for the inputs by itself.
```{r}
plot(x=mtcars$hp, y=mtcars$mpg)
```

However, we could change the parameter **pch** to change the dot. To give the filled dots, pch = 19 will work.
```{r}
plot(x=mtcars$hp, y=mtcars$mpg,
     pch=19)
```

For plotting, using formula also works. 
```{r}
plot(hp ~ mpg, data = mtcars,
     pch=19)
```

The advantage of using the formula is that we do not have to specify the name of axis because the function will take the name from the formula.

What if we would like to change the color? We could give the **col** parametre. It will give the color based on the gear information then!
```{r}
plot(hp ~ mpg, data = mtcars,
     pch=19, col = gear)
```

## Summarization of dataset

### Average

What if we would like to get the average hp for cars with different gears? *mean()* function will be helpful.
```{r}
mean(mtcars$hp)
```

With this command, it will calculate the average value of hp for whole dataset. However, to get what we want, we should subset the dataset.
```{r}
# Creating logical vectors
is.gearEq3 <- mtcars$gear == 3
length(is.gearEq3)
```

```{r}
mtcars.gear3 <- mtcars[is.gearEq3, ]``
dim(mtcars.gear3)
```

Now we have the dataset much smaller than original one, and easy to be used for what we want to get.

Now we will make an average.
```{r}
mean(mtcars.gear3$hp)
```

## Split-apply-combine

Most of analysis will work under the groups. Let's think about the boxplot.
```{r}
boxplot(mpg ~ gear, data=mtcars)
```

It will split the gear information, apply plotting method and combind the informations on the plot! 

One of the most important function is *aggregate()*,  it splits the data into subsets, compute summary statistics for each, and returns the result in a convenient format, in other words, performing combination.

Here, the data will be splitted based on the gear information and apply *mean()* function on it. It will return the average value of hp for each gear information.
```{r}
aggregate(formula = hp ~ gear, data = mtcars, FUN = mean)
```

Well, what if we would like to get standard deviation, not mean?
```{r}
aggregate(formula = hp ~ gear, data = mtcars, FUN = sd)
```

## Sorting

Also we could **sort** dataset and keep the sorted data into environment. This could be worked with *dplyr* but it is also available with basic commands.

*order* will take a position of vectors into ascending order.
```{r}
order(mtcars$hp)
```

So we could use this result to re-structure the dataset and re-arrange the dataset.
```{r}
mtcars[order(mtcars$hp), ]
```

What if we would like to sort the gear information in a manner of ascending of carb?
```{r}
mtcars[order(mtcars$gear, mtcars$carb), ]
```

For descending, you could add minus sign (-) to the values.
```{r}
mtcars[order(mtcars$gear, -mtcars$carb), ]
```

# Tidyverse

Why should we take care of **Tidyverse**? It is already highly structured and usability is better. Not only the code, but **grammer** for coding matters a lot! Also the readability of codes are obviously clear.

For tidyverse, regardless of the function, the first input should be dataset which we would like to put into the function.

## Verbs

Main five verbs provided by dplyr would be *arrange()*, *select()*, *filter()*, *summarise()* and *mutate()*. Also, sometimes *glimpse()* is also nice.

### arrange()

For *arrange()*, you should clarify the dataset, what to be arranged, the order for arrangement in the same group.
```{r}
arrange(mtcars, gear, desc(carb)) # As a descending manner
```

### select()

This verb will be used to select some dataset. In other words, it is useful for subsetting of columns.
```{r}
select(mtcars, hp, mpg)
```

Also the **negative seletion** is possible.
```{r}
select(mtcars, -hp)
```

### filter()

```{r}
filter(mtcars, gear == 3)
```

```{r}
filter(mtcars, gear == 3, hp > 200)
```

### mutate()

When we want to make any change on the column, especially adding new columns, we could use *mutate()* function.

Here as a example, we will put the ratio of mpg to hp as a new column. The new column wîll be added on the very end of the data.
```{r}
mutate(mtcars, mpg_hp_dplyr = mpg/hp)
```

Also the *mutate()* works for several addition.
```{r}
mutate(mtcars, 
       mpg_hp_dplyr = mpg/hp,
       mpg_wt = mpg/wt)
```

### summarise()

It is similar to split-apply-combination concept, especially when working with *group_by()* manner. The name of output and operation should be specified

```{r}
mtcars %>% 
  group_by(gear) %>% 
  summarise(meanHP = mean(hp))
```

Not like *aggregate()*, we do not call the *summarise()* several times.
```{r}
mtcars %>% 
  group_by(gear) %>% 
  summarise(meanHP = mean(hp),
            sdHP = sd(hp))
```

The another version is available, called *summarise_all()*. What if we would like to get all the average, but still consider the group? Just specify which function you would like to run as a input of *summarise_all()*.
```{r}
mtcars %>% 
  group_by(gear) %>% 
  summarise_all(mean)
```


### glimpse()

*glimpse()* is more or less similar to *str()*, it shows all the information about the dataset.
```{r}
glimpse(mtcars)
```

## Pipe operator

Pipe operator could connect the two different commands. It could be typed by shift + command + m and the format is %>% .

When the function is given with pipe operator, the rond bracket () is not required anymore.
```{r}
mtcars$hp %>% log10
```

```{r}
mtcars$hp %>% log10 %>% sin
```

### Example

#### Step by step
```{r}
mtcars.new <- mtcars
mtcars.new$carname <- rownames(mtcars)
head(mtcars.new)
```

```{r}
mazda.cars <- filter(mtcars.new, carname == "Mazda RX4" )
```

```{r}
mazda.cars <- select(mazda.cars, hp, mpg)
mazda.cars
```

#### using pipe operator

```{r}
mtcars.new %>% 
  filter(carname == "Mazda RX4") %>% 
  select(hp, mpg)
```

Q. hp and mpg, but for cars with gear 3 and wt > 2 and sort it as a descending manner of mpg?
```{r}
mtcars %>% 
  filter(gear == 3) %>% 
  filter(wt > 2) %>% 
  select(hp, mpg) %>%
  arrange(hp, desc(mpg))
```

Q. Cars with hp/mpg greater than 9 + what are the gears + make a table and pie chart
```{r}
mtcars %>% 
  mutate(hp_mpg = hp / mpg) %>% 
  filter(hp_mpg > 9) %>% 
  select(gear) %>% 
  table
```

```{r}
mtcars %>% 
  mutate(hp_mpg = hp / mpg) %>% 
  filter(hp_mpg > 9) %>% 
  select(gear) %>% 
  table %>% 
  pie
```

For the dataset generated by using several pipes, it is also available to save it as a new variable. There are two ways of saving.
```{r}
mtcars %>% 
  filter(gear == 3) %>% 
  filter(wt > 2) %>% 
  select(hp, mpg) %>%
  arrange(hp, desc(mpg)) -> data.final
```

```{r}
data.final <- mtcars %>% 
  filter(gear == 3) %>% 
  filter(wt > 2) %>% 
  select(hp, mpg) %>%
  arrange(hp, desc(mpg))
```

