---
title: "R coursework with Sumeet"
subtitle: "4th : File I/O and tidy transformation"
author: "Yura Song"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output:
  html_notebook:
    toc: yes
---

```{r setup, include=FALSE, cache=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = ".")
```

# Preparation

## Installation of required packages

```{r}
install.packages(c("readr", "readxl")) # for reading files
install.packages(c("tibble", "tidyr")) # for changing to tidy data
install.packages("stringr") # for string manipulation
```

## Attach the libraries

If the libraries are installed properly, all libraries should load without any error.

```{r, warning=F, message=F}
library(dplyr)
library(readr)
library(readxl)
library(tibble)
library(tidyr)
library(stringr)
```

# Introduction

In this notebook, as an introduction of R, I will explain to you how to import files on the R global environment and play with the data. Since R is mainly working with *function*, which has same concept with *command* in DOS or LINUX. We will start to learn how to use the functions to play on R.

# Dealing with the Dataset

### Working Directory and files
Before starting the work on R, what we should check is current **location**, where you are working in the system. On R, the working directory should be same as where you put your files. If the current directory and the location of file is different, the system may show the error message.

To check my current directory, we could use *getwd* function, which means get working directory.
```{r}
getwd()
```

However, if your current location is not where you put your files? Then you could use *setwd* function, which means set working directory. In this case, you should put the **address of the target folder**.
```{r}
setwd("~/Desktop/R_course/Notebook/4th - 13April2020/")
```

So, now we are in the directory where we put the files. If you want to check which file is included in the directory which has been set as working directory, you could use *list.files* function.
```{r}
list.files()
```

# Import the files

## read.r

The goal of *readr* is to provide a fast and friendly way to read rectangular data, such as tidy dataset. It is designed to flexibly parse many types of data found in the wild, while still cleanly failing when data unexpectedly changes.

```{r}
TwoDrugs <- read_csv("Session4 Files/ExperimentTwoDrugs.csv")
```

```{r}
head(TwoDrugs)
```

## read.delim

The *read.delim* function is typically used to read in **delimited text files**, where data is organized in a data matrix with rows representing cases and columns representing variables. We can also write a matrix or data frame to a text file using the write.table function. 

When you want to import the data in R, we could use *read.delim* or *read.table* function. Both does not have that difference, but *read.table* only reads the data with table format, as like excel sheet. However, in case of *read.delim*, it reads all the files which is deliminated by something, such as comma or tab. 

The usage of function is as below,

**read.delim(file, header=TRUE, sep="\t")**
  - file – A file location.
  - header – Whether the first line describes the column names.
  - sep – The table delimiter, often times a tab (\t) or comma.

## readxl

Actually, most of you will use **excel files**, rather than .csv or .txt. On R, reading data from an Excel file or Workbook object into a data.frame is also available. 

The **readxl** makes it easy to get data out of Excel and into R. Compared to many of the existing packages (e.g. gdata, xlsx, xlsReadWrite) readxl has no external dependencies, so it’s easy to install and use on all operating systems. It is designed to work with tabular data.

readxl supports both the legacy .xls format and the modern xml-based .xlsx format. 

```{r}
TwoDrugs <- read_xls(path = "Session4 Files/ExperimentTwoDrugs.xls")
head(TwoDrugs)
```

Also, **readxl** allows us to import the different sheet of excel worksheet. To clarify which datasheet we would like to import, we should give *sheet* information.

```{r}
Experiment <- read_xls(path = "Session4 Files/ExperimentWithFeatures.xls", sheet = 1)

Experiment2 <- read_xls(path = "Session4 Files/ExperimentWithFeatures.xls", sheet = 2)
```

What if we know the name of different worksheet? Still the name could be used for giving sheet information.
```{r}
Experiment2 <- read_xls(path = "Session4 Files/ExperimentWithFeatures.xls", sheet = "BadNaming")
```

# Export the files

## write.table

The R base function *write.table()* can be used to export a data frame or a matrix to a file.

```{r}
write.table(Experiment, "experiment.csv", 
            quote=F, sep=",",
            row.names = TRUE, col.names = TRUE)
```

For the *sep* option, we could give tab (\t), semi-colon (;), comma(,) or just a space ( ). It depends in the environment which you would like to use further. For the tab-deliminated, it is recommended to save the data as **.txt** format, rather than **.csv** format.

## **TIP for this session - regarding the file format**

**.csv file** is the one save the data with comma-separated version, in other words, value in each column has been separated in comma (,). For .csv files, you could use normal text editor to open and edit the files. What you should consider is that .csv format only saves the worksheet which you worked at the moment of saving.

Actually, space is not that good separator, since people tends to put the spaces in the name of values. Therefore, it is highly recommended to use other seperators, for instance, comma(,) or tab(\t). _Some European countries, comma could be uses as a decimal point, R would recognize the decimal point as a deliminator_.

When you saved or load the files with the duplicated rownames, it will not let you do further works, just returning the error message that your file has the **duplicated row names**. In this case, you should change the duplicated rownames as A and A.1. 

## saveRDS

There is another way of saving the data, as the **r data structure**. You do not have to specify any further parametres, just object and file name.
```{r}
saveRDS(object = Experiment2, file = "Session4 Files/New_file_saveRDS.rds")
```

```{r}
list.files()
```

# Tidy Transformation

## Load the data

We are going to focus on the excel files which we have worked. We will load the files on the environment.

Before starting the further job, we will clean the environment to initialize. This only removes the variables which we have saved the data, not the libraries.
```{r}
rm(list=ls())
```

```{r}
Expe <- read_xls(path = "Session4 Files/ExperimentTwoDrugs.xls")
glimpse(Expe)
```

Here, as you could see, *NA* values are coming out. It means that it does not have the value, but other columns have. We could check the loaded data with *glimpse()*, but we could not do something further. Therefore, if we would like to do something further, such as transformation. 

## Transformation

Function *gather()*, it gathers all the information and put it as one column. However, we will not use *gather()* function today since it has been retired. To transform, we will use *pivot_longer()*, which has been recommended on the help page of *gather()*. We will use **tidyverse** to work.

Here, we would like to collect all the information and make it into one column. We should specify **data**, **columns** to combine together and **names_to**, the column name which will be given on the transformed data.
```{r}
Expe %>% 
  pivot_longer(cols = everything(), # take all the columns in the data
               names_to = "Treatment",
               values_to = "Cells") # The values to get
```

```{r}
Expe %>% 
  pivot_longer(cols = everything(), 
               names_to = "Treatment",
               values_to = "Cells",
               values_drop_na = TRUE) # Filter out the missing values 
```

However, still the data looks a little bit shuffled. We will arrange the data.

```{r}
Expe %>% 
  pivot_longer(cols = everything(), 
               names_to = "Treatment",
               values_to = "Cells",
               values_drop_na = TRUE) %>% 
  arrange(Treatment) -> Expe.tidy

Expe.tidy
```

Now, all the observations from the same treatment group has been coming up together. All arragned without missing values. 

What if we would like to exclude certain condition? Then we could use negative collection. Or you could specify the names of the certain columns which you would like to transform.
```{r}
Expe %>% 
  pivot_longer(cols = -Drug1, 
               names_to = "Treatment",
               values_to = "Cells",
               values_drop_na = TRUE) %>% 
  arrange(Treatment)
```

## Statistical analysis

Since our data has been transformed as tidy format, we now could run the statistical analysis. 

### Summary statistics

#### Average values

```{r}
Expe.tidy %>% 
  group_by(Treatment) %>%
  summarise(Mean.cells = mean(Cells))
```

#### Standard Deviation (SD)

```{r}
Expe.tidy %>% 
  group_by(Treatment) %>%
  summarise(sd.cells = sd(Cells))
```

#### Standard error of mean (SEM)

We also could calculate SEM, since SEM is the value which SD divided by the square root of number of observations.

Here, we have SD and number of cells.
```{r}
Expe.tidy %>% 
  group_by(Treatment) %>%
  summarise(sd.cells = sd(Cells), 
            Number.Cells=n()) -> Expe.stats # return number of observations following the classification

Expe.stats
```

Now, we are ready to calculate SEM and we will use *mutate()* function.
```{r}
Expe.stats %>% 
  mutate(SEM.cells = sd.cells / sqrt(Number.Cells)) -> Expe.sem 
Expe.sem
```

##### Write own function

What if you would like to do it faster? It is possible by defining the function. Here, we will write the **custom function** by ourselves.

The format of custom function is as below.

**name <- function(argument){operation**
**return()}**.

We could save this function as R file, **.R** format.
```{r}
sem <- function(data, multiply=1){
  sd.data <- sd(data) # calculate stdev
  number.samples <- length(data)
  sem.data <- sd.data/sqrt(number.samples)
  return(sem.data * multiply)
} 
```

When you saved the function as .R format? We could load the function with *source()* function.
```{r}
source("SEM.R")
```

You could run the defined whole structure on the console, then the function will be loaded on your environment. Also, we could use the function as like this.

```{r}
sem(c(1:10))
```

Let's test whether the result is correct. The result should be same if the function has been correctly defined.
```{r}
sd(c(1:10)) / sqrt(10)
```

```{r}
sem(c(1:10), multiply = 2) # Clatification of multiply == SEM * (multiply)
```

Okay, we will use this custom function on our data.
```{r}
Expe.tidy %>% group_by(Treatment) %>%
  summarise(SEM.cells = sem(Cells))
```

#### Coefficient of variation (CV)

CV is the value to see how much the dataset is sparsed.

```{r}
Expe.tidy %>% group_by(Treatment) %>%
  summarise(Mean.cells = mean(Cells), SD.cells = sd(Cells)) %>%
  mutate(CV.cells = SD.cells / Mean.cells)
```

##### Make own function

```{r}
cv <- function(data){
  sd.data <- sd(data)
  mean.data <- mean(data())
  cv.data <- sd.data / mean.data
  return(cv.data)
} 
```

The function is also saved **SEM.R** file, so we need to re-load the file. It is recommended to save the functions as a collection. 
```{r}
source("SEM.R")
```

```{r}
Expe.tidy %>% group_by(Treatment) %>% summarise(CV.cells = cv(Cells))
```

### ANOVA

```{r}
Exp.model <- aov(Cells ~ Treatment, data=Experiment)

# Combined effect of treatment and age
Exp.model1 <- aov(Cell ~ Treatment + Age, data=Experiment) 

# Interaction added
Exp.model2 <- aov(Cell ~ Treatment * Age, data=Experiment)
```

#### Rule out the parametres

```{r}
anova(Exp.model, Exp.model1)
```

```{r}
summary(Exp.model2)
```

# Example - practice

## Example 1

This excel contains linked data, different parametres from two different groups. The parametres are all linked, not separated. **We should not shuffle the data.**
```{r}
Experiment <- read_xls("Session4 Files/ExperimentWithFeatures.xls", sheet=1)

glimpse(Experiment)
```

```{r}
head(Experiment)
```

```{r}
Experiment %>% pivot_longer(cols = -Day, 
                            names_to = c(".value", "Treatment"),
                            names_sep = "_") 
```

```{r}
Experiment %>% 
  pivot_longer(cols = -Day, 
               names_to = c(".value", "Treatment"),
               names_sep = "_") %>%
  drop_na() %>%
  arrange(Treatment)
```

## Example 2

### Rename the column

```{r}
Experiment2 <- read_xls("Session4 Files/ExperimentWithFeatures.xls", sheet=2)
```

```{r}
Experiment2 <- rename(.data = Experiment2,
                      Cells.control = Control,
                      Age.control = Age...2,
                      Gender.control = Gender...3,
                      Cells.drug = Drug,
                      Age.drug = Age...5,
                      Gender.drug = Gender...6)
Experiment2
```

```{r}
Experiment2 %>% 
  pivot_longer(cols = -Day,
               names_to = c(".value","Treatment"),
               names_sep="_") # names_sep does not work with ".".
```
